---
title: "RPres chunks codes"
author: "Joel Le Lay"
date: "2016/04/08"
output: html_document
---
**PAGE 3 CHUNK FOR PLOTS**  

```{r PLOTS_CODE, message=FALSE}
library(fpp) # for  tslm

# Loading data  
s <- readRDS("data/retail_turnover.rds")

# start of the period  to forecast:
r1 <- 1989.75

# end of the period to forecast:
r2 <-  1991.5

# start of the training set:
r <- 1983

# window of quarters used as training set
s1 <- window(s, start=r, end=r1-0.1)

# window of quarters to forecast
s2 <-window(s, start=r1, end=r2-0.1) 

# Number of Quarters to forecast
h <- 4*(r2- r1 ) 

# window of quarters used as training and test sets
s1s2 <-window(s,  start=r, end=r2-0.1) 

# Forecasts (and prediction intervals) made using the following 7 methods   

FUN_fc <- function(s1,h){ 
        
        regFit1 <- tslm(s1 ~ trend)
        regFit2 <- tslm(s1 ~ trend + season) 
        etsFit <- ets(s1)
        
        f1 <- forecast(etsFit, h=h)
        f2 <- snaive(s1, h)  
        f3 <- forecast( regFit2, h=h )
        f4 <- forecast( regFit1, h=h )
        f5 <- meanf(s1, h)
        f6 <- naive(s1, h) 
        f7 <- rwf(s1, h, drift=TRUE)
        
        
        list(Exponential_Smoothing=f1, Seasonal_naive=f2, Multiple_regression =f3, 
             Simple_regression=f4, Mean=f5, Naive=f6, Drift=f7
        )
} 

# getting the forecasts relative to the actual s1 and h
fc <- FUN_fc(s1,h)

# ENTIRE PLOT OR ZOOM                
                
FUN_PZ <- function(sr, fc, type){
        # sr as series (s1s2 or s2), fc as list of 7 forecasts
        # type as "PLOT" or "ZOOM"
        cols <- c("purple","orange",  "red", "green", 
                  "blue", "gray50", "goldenrod3", "indianred4")
                        
        plot(sr, main=ifelse(type=="PLOT", "Training set, test set and forecasts",
                             "Zoom on test set and forecasts"), 
             xlab="Quarters", ylab="", lwd=2, col=cols[1])
        
        for(k in 1:7){
                
                lines(fc[[k]]$mean, lwd=2, col=cols[k+1])
        }
        
        legend(ifelse(type=="PLOT","topleft","topright"), lwd=4, col=cols,
               
               legend=c("Actual values","Exponential smoothing","Seasonal naive",
                        "Multiple regression","Simple regression","Mean", "Naive",
                        "Drift" ))
}

```


```{r PLOT}
FUN_PZ(s1s2, fc,  "PLOT")
```

```{r ZOOM}
FUN_PZ(s2, fc,  "ZOOM")
```

**PAGE 4 CHUNK FOR ACCURACY**  
                
```{r ACCURACY}
# measure of the forecast accuracy: "MAPE" (MEAN ABSOLUTE PERCENTAGE ERROR)
FUNmeasure <- function(f){
        
        round( accuracy(f, s2)[2, "MAPE"]
               ,1 )
}

# MAPE of each forecasting method
a <- sapply(fc, FUNmeasure)

# rank of each method according to its MAPE
rk <- rank(a, ties.method = "min")

aRk <-  data.frame( rbind( a, rk ))
# shortening the columns names so that they are displayed on one line
colnames(aRk) <- c("Exp_smooth", "Seas_naive" ,"Mult_reg", "Simp_reg", "Mean", 
                   "Naive", "Drift")
rownames(aRk)<- c("MAPE", "rank")   

knitr::kable(aRk,padding=0)
```

**PAGE 4 CHUNK FOR TABLE**  

```{r TABLE}
df <- round( data.frame(Actual_values=s2,
                        Exponential_Smoothing=fc[[1]]$mean,
                        Seasonal_naive=fc[[2]]$mean,
                        Multiple_regression=fc[[3]]$mean,
                        Simple_regression=fc[[4]]$mean,
                        Mean=fc[[5]]$mean,
                        Naive=fc[[6]]$mean,
                        Drift=fc[[7]]$mean)
             )

# placing the variables of df in increasing order of their MAPE (in vector "a")

# (note from HELP in Details for "order":any unresolved ties will be left
# in their original ordering)

# first, numbering them with their column number in df 
# (same number as in the vector a, plus 1)
nb <- 2:8
ord <- order( a ) 
nb <- nb[ord]
# leaving the first column at the first place
df <- df[, c(1,nb) ] 
# keeping the best 3 methods only (and the first column)
df <- df[, 1:4 ] 
tdf <- t(df)

knitr::kable(tdf,padding=0)                
```

**PAGE 5 CHUNK FOR STATISTICS**  

To see the code that gave the file statistics.rds, 
see tab STATISTICS / statistics code of the app
and STATISTICS / algorithm logic for the way the code is built.  
(or the files statistics_code.Rmd and algo_logic.txt on GitHUb at the link
https://github.com/lelayj/DATAPRODUCT)   

The result `statistics` of the execution of this code has been saved 
to the file `statistics.rds` due to the time needed for the  execution.

elapsed time:   
3596.872 + 1175.363 + 620.229 + 426.817 + 386.673 + 58.725 + 14.575 = 6279.254  
(Intel® Core™ i5-3210M CPU @ 2.50GHz × 4)  

regarding retail_turnover:  

user  |   system  |    elapsed 
------|---------|-----------
58.768  |     0.000   |   58.725

 

```{r STATISTICS}
statistics <- readRDS("data/statistics.rds")
st6 <- statistics[[6]]
t6 <- t(st6)
colnames(t6) <- c("Exp_Smooth", "Seas_naive" ,"Mult_reg", "Simp_reg", "Mean",
                  "Naive", "Drift")
rownames(t6)<- c("rank1_percent", "MAPE_mean", "MAPE_sd")   
knitr::kable(t6, padding=0)                     
```

